Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
library(ggplot2)
pf = read.csv('pseudo_facebook.tsv', sep='\t')

ggplot(aes(x=gender, y= age),
       data= subset(pf, !is.na(gender))) + geom_boxplot()+
  stat_summary(fun.y=mean, geom="point", shape=4)
ggplot(aes(x=age, y=friend_count),
       data= subset(pf, !is.na(gender)))+
  geom_line(aes(color=gender), stat="summary", fun.y=median)



```
## create a new data frame called 'pf.fc_by_age_gender'
```{r}
library(dplyr)

age_gender = group_by(subset(pf, !is.na(gender)),age, gender)
  pf.fc_by_age_gender = summarise( age_gender, mean_friend_count =mean(friend_count),
             median_friend_count=median(friend_count),
                               n=n())
pf.fc_by_age_gender =ungroup(pf.fc_by_age_gender) 
pf.fc_by_age_gender =arrange(pf.fc_by_age_gender, age)
head(pf.fc_by_age_gender)
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
ggplot(aes(x=age, y=median_friend_count),
       data=pf.fc_by_age_gender)+
  geom_line(aes(color=gender))
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
install.packages('reshape2')
library(reshape2)
pf.fc_by_age_gender.wide =dcast(pf.fc_by_age_gender,
                                age ~ gender,
                                value.var="median_friend_count")
head(pf.fc_by_age_gender.wide)
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
ggplot(aes(x=age, y= female/male),
       data=pf.fc_by_age_gender.wide)+
  geom_line()+
  geom_hline(yintercept = 1, alpha=0.3, linetype = 2)
```

***

### Third Quantitative Variable
Notes:
create new variable in data 
```{r Third Quantitative Variable}

pf$year_joined = 2014-pf$tenure/365
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
summary(pf$year_joined)
pf$year_joined.bucket = cut(pf$year_joined, c(2204, 2009, 2011, 2012, 2014))
head(pf$year_joined.bucket)
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}

```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot(aes(x=age, y=friend_count),
       data=subset(pf, !is.na(year_joined.bucket)))+
  geom_line(aes(color=year_joined.bucket), stat="summary", fun.y=median )
  
```

***

### Friending Rate
Notes:

```{r Friending Rate}
ggplot(aes(x=age, y=friend_count),
       data=subset(pf, !is.na(year_joined.bucket)))+
  geom_line(aes(color=year_joined.bucket), stat="summary", fun.y=median)+
  geom_line(stat= "summary",fun.y=mean, linetype=2 )
```

***

### Friendships Initiated
Notes:
##friending ratio solution
```{r}
with(subset(pf, tenure >= 1), summary(friend_count/tenure))
```


What is the median friend rate? .2205

What is the maximum friend rate? 417

```{r Friendships Initiated}
ggplot(aes(x=tenure, y=friend_initiated/tenure,
       data=subset(pf, tenure >= 1)))+
  geom_line(aes(color=year_joined.bucket))
         
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_smooth(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_smooth(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_smooth(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_smooth(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:
```{r}
yogurt =read.csv("yogurt.csv")
str(yogurt)
yogurt$id =factor(yogurt$id)
```

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
qplot(data=yogurt, x=price, fill="#F79420", binwidth=10)
```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
yogurt = transform(yogurt, all.purchases = strawberry + blueberry + pina.colada + plain+mixed.berry)
summary(yogurt$all.purchases)
```

***

### Prices over Time
Notes:

```{r Prices over Time}
ggplot(aes(x=time, y=price),
       data=yogurt)+
  geom_jitter(alpha =1/4, shape=21)+
  geom_point()
```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(4230)
sample.ids= sample(levels(yogurt$id), 16)

ggplot(aes(x=time, y=price),
       data=subset(yogurt, id %in% sample.ids))+
  facet_wrap(~id)+
  geom_line()+
  geom_point(aes(size =all.purchases), pch=1)
```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:
```{r}
library(GGally)
theme_set(theme_minimal(20))

#set the seed for reprducible results
set.seed(1836)
pf_subset =pf[,c(2:15)]
names(pf_subset)
ggpairs(pf_subset[sample.int(nrow(pf_subset),1000),])
```

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

